---
title: "Sourcing Analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
library(tidyverse)
library(lubridate)
library(plotly)
library(purrr)

library(furrr)
future::plan(multiprocess)

theme_set(theme_light())

set.seed(1984)

normalize <- function(x){
  (x-mean(x))/sd(x)
}

fit_normalize <- function(x, ref) {
  (x - mean(ref)) / sd(ref)
}

reverse_normalize <- function(x,ref){
  x * sd(ref) + mean(ref)
}
```

## Import Data
```{r}

df <- rio::import("qa-0854x.xlsx", sheet = "TV Daily Sales Data", skip=6, n_max = 418) %>% 
  janitor::clean_names() %>% 
  mutate(date = date %>% as.Date()) # original column was parsed as datetime, contains precision down to millisec, cause problem downstream
```

## Re-Create Chart
```{r}
ggplotly(df %>% 
  ggplot(aes(date, units_sold, color = as.factor(lec_id)))+
  geom_line()+
  labs(title = "Sales of Collections",
       x = "",
       y= "",
       color = "Collections")
  )
```

## Transform data to collection level for next collection prediction
```{r}
by_collection <- df %>% 
  group_by(lec_id) %>% 
  group_modify(function(.x,.y){
    start <- first(.x$date)
    end <- last(.x$date)
    duration <- as.integer(end-start)
    weekends <- sum(weekdays(.x$date) %in% c("Saturday", "Sunday"))

    
    
    data.frame(sales = sum(.x$units_sold),
               start = start,
               end   = end,
               duration = duration,
               weekends = weekends
               )
  }) %>% 
  ungroup() %>% 
  mutate(holiday = if_else(lec_id %in% c(8,17:19), TRUE, FALSE))

by_collection
```

## Quick look at distribution of sales
```{r}
by_collection %>% 
  ggplot(aes(sales)) +
  geom_histogram()
```

## Model log(sales) with available features and check fit
```{r}
by_collection %>% 
  select(-lec_id,-start,-end) %>% 
  lm(log(sales) ~ duration + holiday , data = .) -> lm_mod # log(sales) model naturally have interactions

lm_mod %>% 
  summary

par(mfrow = c(2, 2))
lm_mod %>% plot(ask=F)
par(mfrow = c(1, 1))
```

## Two ways of generating simulated demands, contrasted

```{r}

N <- 100000

mean_pred <- predict(lm_mod, data.frame(duration = 21, holiday = F))

lec_sigma <- sigma(lm_mod) # could use this for lognormal distributions

demands <- exp(mean_pred+sample(lm_mod$residuals,N, replace = T)) %>% round() # sample was too small, recycling the residual instead of using random distributions

demands2 <- rnorm(N, mean_pred, lec_sigma) %>% exp %>% round()
```

```{r}
tibble(resid = demands, distribution = demands2) %>% 
  gather() %>% 
  ggplot(aes(value, col = key))+
  geom_freqpoly()+
  labs(title = "Comparing three ways of generating demand")
```

## Calculate profit based on quantity level:350
```{r}
Q <- 350

price <- 135
cost <- 65

sims <- tibble(demand = demands2, 
       sales = if_else(demand<=Q, demand, Q),
       revenue = sales * price,
       profit = revenue - Q * cost
       )

sims %>% 
  ggplot(aes(profit))+
  geom_histogram()+
  labs(title =glue::glue("Profit distribution when purchase is {Q}"))

summary(sims$profit)
```

## Generate profit simulation to arbitrary levels

```{r}
expected_profit <- function(Q, Ds = demands2){
  price <- 135
  cost <- 65
  
  sims <- tibble(demand = Ds, 
         sales = if_else(demand<=Q, demand, Q),
         revenue = sales * price,
         profit = revenue - Q * cost
         ) 

  data.frame(profit_expected = mean(sims$profit), profit_sd = sd(sims$profit))
}
```
```{r}
expected_profit(350) # test
```

## Optimal sourcing quantity: outsourcing only
```{r}
diff_Qs <- tibble(
      Qs = c(200,201:350),
      e_profit = map(Qs, expected_profit)) %>% 
    unnest(e_profit)

optimal_Q <- diff_Qs %>% 
  filter(profit_expected == max(profit_expected)) %>% 
  pull(Qs)

max_profit_outsourcing <- diff_Qs %>% 
  filter(profit_expected == max(profit_expected)) %>% 
  pull(profit_expected)

ggplotly(       
       diff_Qs %>% 
  ggplot(aes(Qs, profit_expected))+
  geom_line()+
  geom_vline(xintercept = optimal_Q, lty=2)+
  scale_y_continuous(labels = scales::dollar)+
  geom_text(data = data.frame(x = optimal_Q, y = max_profit_outsourcing, label = glue::glue("Source {optimal_Q} from Ukraine proves to be most profitable")),mapping = aes(x = x, y = y, label = label), size = 3.9, alpha = 1, inherit.aes = FALSE, check_overlap = T, nudge_x = 40)+

  labs(title = "Expected profit under different sourcing quantity",
       x="",
       y=""
       )
  )
```
```{r}
optimal_Q <- diff_Qs %>% 
  filter(profit_expected == max(profit_expected)) %>% 
  pull(Qs)

max_profit_outsourcing <- diff_Qs %>% 
  filter(profit_expected == max(profit_expected)) %>% 
  pull(profit_expected)

```


Source `r optimal_Q` from Ukraine proves to be most profitable, with profit `r `max_profit_outsourcing`

# Local Sourcing

```{r}
Q <- 253

price <- 135
cost_out <- 65
cost_in <- 95

sims <- tibble(demand = demands2, 
       local_needed = demand>Q,  
       sales_out = if_else(!local_needed, demand, Q),
       sales_in = if_else(local_needed,
                          if_else(demand<=Q+(21-7)*12,demand-Q, (21-7)*12),
                          0),
       revenue = (sales_in + sales_out) * price,
       profit = revenue - Q * cost - sales_in *cost_in
       )

sims %>% 
  ggplot(aes(profit))+
  geom_histogram()+
  labs(title =glue::glue("Profit distribution when purchase is {Q}"))

summary(sims$profit)
```


```{r}
expected_profit_local <- function(Q, Ds = demands2, duration = 21){
  price <- 135
  cost_in <- 95
  cost_out <- 65
  
  sims <- tibble(demand = Ds, 
         local_needed = demand>Q,  
         sales_out = if_else(!local_needed, demand, Q),
         sales_in = if_else(local_needed,
                            if_else(demand<=Q+(duration-7)*12,demand-Q, (duration-7)*12),
                            0),
         revenue = (sales_in + sales_out) * price,
         profit = revenue - Q * cost_out - sales_in *cost_in
         )

  mean(sims$profit)
}
```

```{r}
expected_profit_local(253) #test
```

```{r}
diff_Qs <- tibble(
      Qs = c(200,seq(201,260)), # turn Qs to double
      e_profit = map_dbl(Qs, expected_profit_local)       
       )

near_optimal_Q <- diff_Qs %>% 
  filter(e_profit == max(e_profit)) %>% 
  pull(Qs)

max_profit_local <- diff_Qs %>% 
  filter(e_profit == max(e_profit)) %>% 
  pull(e_profit)



ggplotly(diff_Qs %>% 
  ggplot(aes(Qs, e_profit))+
  geom_line()+
  geom_vline(xintercept = near_optimal_Q, lty=2)+
  scale_y_continuous(labels = scales::dollar)+
  geom_text(data = data.frame(x = near_optimal_Q, y = max_profit_local, label = glue::glue("Source {near_optimal_Q} from Ukraine proves to be most profitable")),mapping = aes(x = x, y = y, label = label), size = 3.9, alpha = 1, inherit.aes = FALSE, nudge_x = 20)+

  labs(title = "Expected profit under different sourcing quantity",
       x="",
       y=""
       )
  )
```


```{r}
max_profit_local - max_profit_outsourcing
```

# Backtesting

## Availability of data

For LEC 11, only LEC data from 1-10 is available.

```{r}
LEC <- 23

# Model demand
by_collection %>%
  filter(lec_id<LEC) %>% # only data prior to this LEC is available
  select(-lec_id,-start,-end) %>% # This is not strictly necessary, just here to not include id as a variable
  lm(log(sales) ~ duration + holiday , data = .) -> lm_mod

actual_demand <- by_collection %>% filter(lec_id == LEC) %>% pull(sales)

# Simulate demands
N <- 100000

mean_pred <- predict(lm_mod, by_collection %>% filter(lec_id==LEC))
# mean_pred <- predict(lm_mod, data.frame(duration = 21, holiday = F)) # only for testing

lec_sigma <- sigma(lm_mod) # could use this for lognormal distributions

demands <- rnorm(N, mean_pred, lec_sigma) %>% exp %>% round()

# Outsource only quantity
range_Qs <- c(2:1500)

profits <- tibble(
      Qs = c(1, range_Qs),
      e_profit = future_map(Qs, ~expected_profit(Q = .x, Ds = demands))) %>% 
    unnest(e_profit)


optimal <- profits %>% filter(profit_expected == max(profit_expected ))

out_optimal_Q <- optimal %>% pull(Qs)
out_optimal_profit <- optimal %>% pull(profit_expected)

out_actual_profit <- {
  sales <- if(actual_demand > out_optimal_Q) out_optimal_Q else actual_demand
  sales * 135 - out_optimal_Q * 65
}

# With near-sourcing options

duration <- by_collection %>% filter(lec_id==LEC) %>% pull(duration)

# duration <- 21 # only for testing

profits_in <- tibble(
  Qs = c(1, range_Qs),
  e_profit = future_map_dbl(Qs, ~expected_profit_local(Q = .x, Ds = demands, duration = duration))
)

optimal <- profits_in %>% filter(e_profit == max(e_profit ))

in_optimal_Q <- optimal %>% pull(Qs)
in_optimal_profit <- optimal %>% pull(e_profit)

in_actual_profit <- {
  local_needed <- actual_demand > in_optimal_Q
  sales_out <- if(!local_needed) actual_demand else in_optimal_Q
  sales_in <- if_else(local_needed,
                      if_else(actual_demand<=in_optimal_Q+(duration-7)*12,
                              actual_demand-in_optimal_Q, 
                              (duration-7)*12),
                            0)
  revenue <-  (sales_in + sales_out) * 135
  profit = revenue - in_optimal_Q * 65 - sales_in * 95
  profit
}


output <- by_collection %>% 
  filter(lec_id == LEC) %>% 
  mutate(ln_mean_demand = mean_pred,
         ln_mean_sigma = lec_sigma,
         sim_demand_mean = mean(demands)
         ) %>% 
  mutate(out_optimal_Q = out_optimal_Q,
         out_optimal_profit = out_optimal_profit,
         out_actual_profit = out_actual_profit,
         in_optimal_Q = in_optimal_Q,
         in_optimal_profit = in_optimal_profit,
         in_actual_profit = in_actual_profit
         )

```



```{r}

backtest <- function(LEC = 23){

# Model demand
by_collection %>%
  filter(lec_id<LEC) %>% # only data prior to this LEC is available
  select(-lec_id,-start,-end) %>% # This is not strictly necessary, just here to not include id as a variable
  lm(log(sales) ~ duration + holiday , data = .) -> lm_mod

actual_demand <- by_collection %>% filter(lec_id == LEC) %>% pull(sales)

# Simulate demands
N <- 100000

mean_pred <- predict(lm_mod, by_collection %>% filter(lec_id==LEC))
# mean_pred <- predict(lm_mod, data.frame(duration = 21, holiday = F)) # only for testing

lec_sigma <- sigma(lm_mod) # could use this for lognormal distributions

demands <- rnorm(N, mean_pred, lec_sigma) %>% exp %>% round()

# Outsource only quantity
range_Qs <- c(2:1500)

profits <- tibble(
      Qs = c(1, range_Qs),
      e_profit = future_map(Qs, ~expected_profit(Q = .x, Ds = demands))) %>% 
    unnest(e_profit)


optimal <- profits %>% filter(profit_expected == max(profit_expected ))

out_optimal_Q <- optimal %>% pull(Qs)
out_optimal_profit <- optimal %>% pull(profit_expected)

out_actual_profit <- {
  sales <- if(actual_demand > out_optimal_Q) out_optimal_Q else actual_demand
  sales * 135 - out_optimal_Q * 65
}

# With near-sourcing options

duration <- by_collection %>% filter(lec_id==LEC) %>% pull(duration)

# duration <- 21 # only for testing

profits_in <- tibble(
  Qs = c(1, range_Qs),
  e_profit = future_map_dbl(Qs, ~expected_profit_local(Q = .x, Ds = demands, duration = duration))
)

optimal <- profits_in %>% filter(e_profit == max(e_profit ))

in_optimal_Q <- optimal %>% pull(Qs)
in_optimal_profit <- optimal %>% pull(e_profit)

in_actual_profit <- {
  local_needed <- actual_demand > in_optimal_Q
  sales_out <- if(!local_needed) actual_demand else in_optimal_Q
  sales_in <- if_else(local_needed,
                      if_else(actual_demand<=in_optimal_Q+(duration-7)*12,
                              actual_demand-in_optimal_Q, 
                              (duration-7)*12),
                            0)
  revenue <-  (sales_in + sales_out) * 135
  profit = revenue - in_optimal_Q * 65 - sales_in * 95
  profit
}


output <- by_collection %>% 
  filter(lec_id == LEC) %>% 
  mutate(ln_mean_demand = mean_pred,
         ln_mean_sigma = lec_sigma,
         sim_demand_mean = mean(demands)
         ) %>% 
  mutate(out_optimal_Q = out_optimal_Q,
         out_optimal_profit = out_optimal_profit,
         out_actual_profit = out_actual_profit,
         in_optimal_Q = in_optimal_Q,
         in_optimal_profit = in_optimal_profit,
         in_actual_profit = in_actual_profit
         )
  output
}

```

```{r}
backtests <- tibble(LEC = c(11,12:23))
```

```{r}

backtest_result <-  future_map_dfr(11:23, backtest)
```

```{r}
backtest_result %>% 
  rio::export("backtest_results.xlsx")
```


