---
title: "Sourcing Analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
library(tidyverse)
library(RcppBDT)
library(lubridate)
library(plotly)
library(purrr)
library(rstan)
library(brms)
theme_set(theme_light())

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores()-1)

set.seed(1984)

normalize <- function(x){
  (x-mean(x))/sd(x)
}

fit_normalize <- function(x, ref) {
  (x - mean(ref)) / sd(ref)
}

reverse_normalize <- function(x,ref){
  x * sd(ref) + mean(ref)
}
```

## Import Data
```{r}

df <- rio::import("qa-0854x.xlsx", sheet = "TV Daily Sales Data", skip=6, n_max = 418) %>% 
  janitor::clean_names() %>% 
  mutate(date = date %>% as.Date()) # original column was parsed as datetime, contains precision down to millisec, cause problem downstream
```

## Re-Create Chart
```{r}
ggplotly(df %>% 
  ggplot(aes(date, units_sold, color = as.factor(lec_id)))+
  geom_line()+
  labs(title = "Sales of Collections",
       x = "",
       y= "",
       color = "Collections")
  )
```

## Holidays for boolean features
```{r}
holidays <- tibble(years = 2000:2050,
                   BFriday = map(years, ~getNthDayOfWeek(fourth, Fri, Nov, .x)),
                   boxing = lubridate::make_date(years,12,26)
                  ) %>% 
                    unnest(BFriday) 
```

## Transform data to collection level for next collection prediction
```{r}
by_collection <- df %>% 
  group_by(lec_id) %>% 
  group_modify(function(.x,.y){
    start <- first(.x$date)
    end <- last(.x$date)
    duration <- as.integer(end-start)
    weekends <- sum(weekdays(.x$date) %in% c("Saturday", "Sunday"))

    have_black_friday <- any(.x$date %in% holidays$BFriday)
    
    have_boxing <- any(.x$date %in% holidays$boxing)
    
    data.frame(sales = sum(.x$units_sold),
               start = start,
               end   = end,
               duration = duration,
               weekends = weekends,
               have_black_friday = have_black_friday,
               have_boxing = have_boxing
               # years = years
               )
  }) 
```

## Quick look at distribution of sales
```{r}
by_collection %>% 
  ggplot(aes(sales)) +
  geom_histogram()
```

## Model log(sales) with available features and check fit
```{r}
by_collection %>% 
  select(-lec_id,-start,-end) %>% 
  lm(log(sales) ~ duration  + weekends + have_black_friday + have_boxing , data = .) -> lm_mod # log(sales) model naturally have interactions

# + have_black_friday:duration + weekends: duration
# + I(duration^2)

lm_mod %>% 
  summary

par(mfrow = c(2, 2))
lm_mod %>% plot(ask=F)
par(mfrow = c(1, 1))
```

## Three ways of generating simulated demands, contrasted

```{r}

N <- 100000

mean_pred <- predict(lm_mod, data.frame(duration = 21, weekends = 6,have_black_friday = F, have_boxing = F))

lec_sigma <- sigma(lm_mod) # could use this for lognormal distributions

demands <- exp(mean_pred+sample(lm_mod$residuals,N, replace = T)) %>% round() # sample was too small, recycling the residual instead of using random distributions

demands2 <- rnorm(N, mean_pred, lec_sigma) %>% exp %>% round()
```

```{r}
df <- by_collection %>% 
  ungroup() %>% 
  select(-lec_id,-start,-end) %>% 
  mutate(log_sales = log(sales)) %>% 
  select(-sales)

df_scaled <- df %>% 
  mutate(across(c(duration,log_sales, weekends), normalize))

iter <- floor(N / 4) * 2 # 4 chains and half goes to warmup

lec_brm <- brm(log_sales ~ duration  + weekends + have_black_friday + have_boxing , data = df_scaled, iter = iter)

newdata <- data.frame(duration = fit_normalize(21, df$duration), 
                      weekends = fit_normalize(6, df$weekends),
                      have_black_friday = F, have_boxing = F)

pred <- predict(lec_brm, newdata = newdata, summary = F) %>% 
  reverse_normalize(df$log_sales) %>% 
  exp %>% 
  round
```

```{r}
tibble(resid = demands, distribution = demands2, bayes = pred) %>% 
  gather() %>% 
  ggplot(aes(value, col = key))+
  geom_freqpoly()+
  labs(title = "Comparing three ways of generating demand")
```

## Calculate profit based on quantity level:350
```{r}
Q <- 350

price <- 135
cost <- 65

sims <- tibble(demand = pred, 
       sales = if_else(demand<=Q, demand, Q),
       revenue = sales * price,
       profit = revenue - Q * cost
       )

sims %>% 
  ggplot(aes(profit))+
  geom_histogram()+
  labs(title =glue::glue("Profit distribution when purchase is {Q}"))

summary(sims$profit)
```

## Generate profit simulation to arbitrary levels

```{r}
expected_profit <- function(Q, Ds = pred){
  price <- 135
  cost <- 65
  
  sims <- tibble(demand = Ds, 
         sales = if_else(demand<=Q, demand, Q),
         revenue = sales * price,
         profit = revenue - Q * cost
         ) 

  data.frame(profit_expected = mean(sims$profit), profit_sd = sd(sims$profit))
}
```
```{r}
expected_profit(350) # test
```

## Optimal sourcing quantity: outsourcing only
```{r}
diff_Qs <- tibble(
      Qs = c(seq(250,310, by=10), 310:350, seq(350,400, by=10)),
      e_profit = map(Qs, expected_profit)) %>% 
    unnest(e_profit)

optimal_Q <- diff_Qs %>% 
  filter(profit_expected == max(profit_expected)) %>% 
  pull(Qs)

max_profit_outsourcing <- diff_Qs %>% 
  filter(profit_expected == max(profit_expected)) %>% 
  pull(profit_expected)

ggplotly(       
       diff_Qs %>% 
  ggplot(aes(Qs, profit_expected))+
  geom_line()+
  geom_vline(xintercept = optimal_Q, lty=2)+
  scale_y_continuous(labels = scales::dollar)+
  geom_text(data = data.frame(x = optimal_Q, y = max_profit_outsourcing, label = glue::glue("Source {optimal_Q} from Ukraine proves to be most profitable")),mapping = aes(x = x, y = y, label = label), size = 3.9, alpha = 1, inherit.aes = FALSE, check_overlap = T, nudge_x = 40)+

  labs(title = "Expected profit under different sourcing quantity",
       x="",
       y=""
       )
  )
```
```{r}
optimal_Q <- diff_Qs %>% 
  filter(profit_expected == max(profit_expected)) %>% 
  pull(Qs)

max_profit_outsourcing <- diff_Qs %>% 
  filter(profit_expected == max(profit_expected)) %>% 
  pull(profit_expected)

```


Source 315 from Ukraine proves to be most profitable

# Local Sourcing

```{r}
Q <- 350

price <- 135
cost_out <- 65
cost_in <- 95

sims <- tibble(demand = demands, 
       local_needed = demand>Q,  
       sales_out = if_else(!local_needed, demand, Q),
       sales_in = if_else(local_needed,
                          if_else(demand<=Q+21*12,demand-Q, 21*12),
                          0),
       revenue = (sales_in + sales_out) * price,
       profit = revenue - Q * cost - sales_in *cost_in
       )

sims %>% 
  ggplot(aes(profit))+
  geom_histogram()+
  labs(title =glue::glue("Profit distribution when purchase is {Q}"))

summary(sims$profit)
```


```{r}
expected_profit_local <- function(Q, Ds = pred){
  price <- 135
  cost_in <- 95
  cost <- 65
  
  sims <- tibble(demand = demands, 
         local_needed = demand>Q,  
         sales_out = if_else(!local_needed, demand, Q),
         sales_in = if_else(local_needed,
                            if_else(demand<=Q+21*12,demand-Q, 21*12),
                            0),
         revenue = (sales_in + sales_out) * price,
         profit = revenue - Q * cost - sales_in *cost_in
         )

  mean(sims$profit)
}
```

```{r}
expected_profit_local(250)
```

```{r}
diff_Qs <- tibble(
      Qs = c(249,seq(250,350)), # turn Qs to double
      e_profit = map_dbl(Qs, expected_profit_local)       
       )
ggplotly(diff_Qs %>% 
  ggplot(aes(Qs, e_profit))+
  geom_line()+
  geom_vline(xintercept = 270, lty=2)+
  scale_y_continuous(labels = scales::dollar)+
  geom_text(data = data.frame(x = 298, y = 19700, label = "Source 270 from Ukraine proves to be most profitable"),mapping = aes(x = x, y = y, label = label), size = 3.86605783866058, alpha = 1, inherit.aes = FALSE)+

  labs(title = "Expected profit under different sourcing quantity",
       x="",
       y=""
       )
  )
```

```{r}
diff_Qs %>% 
  filter(e_profit == max(e_profit))

max_profit_local <- diff_Qs %>% 
  filter(e_profit == max(e_profit)) %>% 
  pull(e_profit)

```

```{r}
max_profit_local - max_profit_outsourcing
```

